---
title: "Investigate Various Exponential Distribution Statistics"
author: "YeongWei"
date: "Saturday, November 21, 2015"
output: pdf_document
---

# Overview

This document aims to investigate the various statistical properties for Exponential Distribution by comparing the sample and theoretical statistical properties.

# Simulations

The R Code below attempts to produce Exponential Distributions of 100, 1000, 10000 number of observations for 1000 number of simulations with $\lambda$ as 0.2 then for each simulation compute the mean and variance. 

```{r echo = TRUE}
set.seed(10000);

lambda <- 0.2;
numberOfSimulations <- 1000;
numberOfObservations <- c(100, 1000, 10000);

statisticsByGroup <- data.frame(Group = c(), Mean = c(), Variance = c());
for (x in numberOfObservations) {
  meanAccumulator <- c();
  varianceAccumulator <- c();
  for (n in 1 : numberOfSimulations) {
    expSampleDistribution <- rexp(x, lambda);
    
    meanAccumulator <- c(meanAccumulator, mean(expSampleDistribution));
    varianceAccumulator <- c(varianceAccumulator, var(expSampleDistribution));
  }
  group <- rep(x, numberOfSimulations);
  groupMeanVariance <- cbind(Group = group, Mean = meanAccumulator, Variance = varianceAccumulator);
  statisticsByGroup <- rbind(statisticsByGroup, groupMeanVariance);
}
```

# Theoretical Statistics

Given that, mean is $\mu$ = 1 / $\lambda$, standard deviation is $\sigma$ = 1 / $\lambda$ and $\lambda$ = 0.2. The following attempts to compute the theoretical mean, standard deviation and variance.

## Mean, Standard Deviation and Variance

```{r echo = TRUE}
mu <- 1 / lambda;
sd <- 1 / lambda;
var <- sd ^ 2;
c(mu, sd, var);
```

# Statistics Inference

This section attempts to investigate the sample mean and variance against the theoretical ones.

```{r echo = FALSE}
library(ggplot2);
```

## Sample Mean Versus Theoretical Mean

```{r}
PLOT <- ggplot(data = statisticsByGroup, aes(x = Mean, fill = Group));
PLOT <- PLOT + geom_histogram(alpha = 0.20, binwidth= 0.3, colour = "black");
PLOT <- PLOT + facet_grid(. ~ Group);
print(PLOT);
```

## Sample Variance Versus Theoretical Variance

```{r}
PLOT <- ggplot(data = statisticsByGroup, aes(x = Variance, fill = Group));
PLOT <- PLOT + geom_histogram(alpha = 0.20, binwidth= 0.3, colour = "black");
PLOT <- PLOT + facet_grid(. ~ Group);
print(PLOT);
```

