---
title: "Investigation on ToothGrowth Dataset"
author: "YeongWei"
date: "Sunday, November 22, 2015"
output: pdf_document
---

# Overview

This document aims to investigate the data associated to the ToothGrowth dataset.

```{r echo = FALSE}
library(ggplot2);
data(ToothGrowth);
```

# Data Exploratory and Analysis
```{r echo = FALSE}
names(ToothGrowth) <- c("Length", "SupplementType", "Dosage");
ToothGrowth$SupplementType <- as.character(ToothGrowth$SupplementType);
ToothGrowth$Dosage <- as.character(ToothGrowth$Dosage);
ToothGrowth$SupplementType[ToothGrowth$SupplementType == "OJ"] <- "Orange Juice";
ToothGrowth$SupplementType[ToothGrowth$SupplementType == "VC"] <- "Vitamin C";
```

## Scatter Plots
```{r echo = FALSE, fig.height = 2.8}
PLOT <- ggplot(data = ToothGrowth, aes(x = Dosage, y = Length));
PLOT <- PLOT + geom_point();
PLOT <- PLOT + facet_wrap(~ SupplementType);
print(PLOT);
```

## Box Plots
```{r echo = FALSE, fig.height = 2.8}
PLOT1 <- ggplot(data = ToothGrowth, aes(x = Dosage, y = Length));
PLOT1 <- PLOT1 + geom_boxplot();
PLOT1 <- PLOT1 + facet_wrap(~ SupplementType);
print(PLOT1);
```

## Observations and Discussions
1. As overview, the higher dosage of supplement, the higher the probability of getting higher length.
2. Vitamin C has the potential of getting higer length than Orange Juice.
3. Orange Juice has a tendency of logistic length growth and Vitamin C has a tendency of linear length growth.

# Analysis on Tooth Growth by Supplement Type and Dosage

## By Supplement Type

### Equal Variance
```{r echo = TRUE}
stats <- t.test(ToothGrowth$Length[ToothGrowth$SupplementType == "Orange Juice"],
       ToothGrowth$Length[ToothGrowth$SupplementType == "Vitamin C"],
       paired = FALSE, var.equal = TRUE);
c(stats$p.value, stats$conf.int, stats$estimate);
```

### Unequal Variance
```{r echo = TRUE}
stats <- t.test(ToothGrowth$Length[ToothGrowth$SupplementType == "Orange Juice"],
       ToothGrowth$Length[ToothGrowth$SupplementType == "Vitamin C"],
       paired = FALSE, var.equal = FALSE);
c(stats$p.value, stats$conf.int, stats$estimate);
```

## By Dosage

## Equal Variance
```{r echo = TRUE}
stats <- t.test(ToothGrowth$Length[ToothGrowth$Dosage == 0.5],
       ToothGrowth$Length[ToothGrowth$Dosage == 2], paired = FALSE, var.equal = TRUE);
c(stats$p.value, stats$conf.int, stats$estimate);
```

## Unequal Variance
```{r echo = TRUE}
stats <- t.test(ToothGrowth$Length[ToothGrowth$Dosage == 0.5],
       ToothGrowth$Length[ToothGrowth$Dosage == 2], paired = FALSE, var.equal = FALSE);
c(stats$p.value, stats$conf.int, stats$estimate);
```

# Summary
1. All test consist of negative values for confidence interval implies a small sample size.
2. There is minimal impact from variance equality assumption.
3. Confidence Interval test by Supplement Type yields P-Value approximates to 0.06 and includes 0 in the Confidence Interval.
4. Confidence Interval test by Dosage yields P-Value approximates to 0 (multiplication with e-14) and does not include 0 in the Confidence Interval.

# Conclusion
1. The Confidence Interval by Supplement Type includes 0 and P-Value large than 0.05 (standard cut-off point). Therefore the Supplement Type does not have apparent contibution to tooth length.
2. The Confidence Interval by Dosage does not include 0 and P-Value close to 0. Therefore the amount Dosage has apparent contribution to tooth length. 
3. Regardless of the Supplement Type, the amount of Dosage has a better impact on tooth growth.
